Staged Abstract Interpreters

### 2019-08-16 paper/10_intro.tex
Fixing 62

However, besides the intrinsic complexity of static analysis, there
are additional abstraction penalties from these high-level
implementation approaches. First, similar to concrete interpreters,
an abstract interpreter analyzes the program by traversing the
abstract syntax tree (AST), which incurs interpretive overhead, such
as the pattern matching on the ASTs and recursive calls on the
subexpressions.
Such overhead can be negligible if the abstract interpreter only runs
on the program a single time, but can accumulate significantly if
run repeatedly (for example, on libraries).  Second, it is common that
abstract interpreters written in pure languages extensively use effect
systems to implement the semantics of abstract interpretation. For
example, the collecting interpreter that returns all possible runtime
values is a form of nondeterminism, where nondeterminism monads can
help.  Although such pure approaches have their own merits, compared
with imperative stateful implementations, they are significantly
slower.

Roughly at the same when abstract interpretation was proposed in
1970s, \citet{futamura1971partial, Futamura1999} revealed a close
connection between interpreters and compilers through a hierarchy of
specializations, namely, Futamura Projections. The first Futamura
projection shows that specializing an interpreter to the input program
removes interpretation overhead and yields an equivalent executable
that runs faster.  The challenge here: can we apply the same idea to
general abstract interpreters to make hem run faster?
To achieve this, we need to handle a nonstandard semantics that
extensively uses nondeterminism, operates on abstract domains, and
computes fixed-point iteratively with a cache.

>>>>

However, besides the intrinsic complexity of static analysis, there are
additional abstraction penalties from these high-level implementation
approaches. First, similar to concrete interpreters, an abstract interpreter
analyzes the program by traversing the abstract syntax tree (AST), which incurs
interpretive overhead, such as the pattern matching on the ASTs and recursive
calls on the subexpressions. Such overhead can be negligible if the abstract
interpreter only runs on the program a single time, but can accumulate
significantly if run repeatedly (for example, on libraries).  Second, it is
common that abstract interpreters written in pure languages extensively use
effect systems to implement the semantics of abstract interpretation. For
example, the collecting interpreter that returns all possible runtime values is
nondeterministic.  Although such pure approaches have their own merits, compared
with imperative, stateful, implementations, they are significantly slower. In
order to reconcile these approaches, one must use a programming technique which
allows for high-level, functional abstractions during implementation, but which
bypasses the incurred overhead at runtime.

Roughly at the same time when abstract interpretation was proposed in
1970s, \citet{futamura1971partial, Futamura1999} revealed a close
connection between interpreters and compilers through a hierarchy of
specializations, namely, Futamura Projections. The first Futamura
projection shows that specializing an interpreter to the input program
removes interpretation overhead and yields an equivalent executable
that runs faster.  The challenge here: can we apply the same idea to
general abstract interpreters to make hem run faster?
To achieve this, we need to handle a nonstandard semantics that
extensively uses nondeterminism, operates on abstract domains, and
computes fixed-point iteratively with a cache.
